
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(ggforce)
library(cowplot)
library(tidyr)
library(lme4)
library(lmerTest)
library(stringr)
library(emmeans)
library(foreach)
library(parallel)
library(doSNOW)
library(purrr)
options(dplyr.summarise.inform = FALSE)
num_multithread_cores <- 5
```

```{r}
datasets <- c("Filler", "ClassicGP", "RelativeClause", 
              "AttachmentAmbiguity", "Agreement")
#all models
model_names <- read.csv('../data/surprisals/all_lm_names.csv')$x
measures <- c('gz', 'gp', 'regout')

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}
```


Fit filler models
```{r}
et_with_surprisal <- readRDS("../data/empirical_reading_data/ET_data_processed_w_surprisals.rds")
et_with_surprisal <- et_with_surprisal %>% subset(!is.na(logfreq_s) & !is.na(logfreq_p1_s) & !is.na(logfreq_p2_s) & (sent_length != region-1))
```

```{r}
###Formulae for all models
formula_lm <- "MEASURE ~ SURPRISAL + SURPRISAL_p1 + 
  region_s + 
  logfreq_s*length_s + 
  logfreq_p1_s*length_p1_s + 
  (1 + SURPRISAL + SURPRISAL_p1 + logfreq_s + logfreq_p1_s || subj) + 
  (1 | item)"

formula_nosurp <- "MEASURE ~ region_s + 
                 logfreq_s*length_s + 
                 logfreq_p1_s*length_p1_s + 
                 (1 + logfreq_s + logfreq_p1_s || subj) + 
                 (1 | item)"

```

Fit filler models (non regression-contingent)
gz = gaze duration
gp = go-past time
regout = first pass regressions out

```{r}

measures <- c('gz', 'gp', 'regout')

num_cores = min(num_multithread_cores, detectCores()-1)
cl <- makeCluster(num_cores)
registerDoSNOW(cl)
all_tasks <- expand.grid(model_name = model_names, measure = measures, stringsAsFactors = FALSE)
total <- nrow(all_tasks)
pb <- txtProgressBar(max = total, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
result <- 
foreach(i = 1:nrow(all_tasks), .packages=c('dplyr', 'stringr', 'lme4'), .options.snow = opts) %dopar% {
  model_name <- all_tasks$model_name[i]
  measure <- all_tasks$measure[i]
    if(model_name == 'baseline_surp'){
      curr_formula <- str_replace_all(formula_nosurp, "MEASURE", measure)
    }
    else{
      curr_formula <- str_replace_all(str_replace_all(formula_lm, "MEASURE", measure), "SURPRISAL", model_name)
    }
    vars_in_model <- c(all.vars(as.formula(curr_formula)), 'dataset', 'cond', 'regressed', 'eos_region', 'bos_region')
    minimal_dataset <- et_with_surprisal %>% select(all_of(vars_in_model)) %>%
                        filter(dataset == "Filler" & !bos_region & !eos_region)
    if(measure ==  'regout'){ #binary measures
      filler.model <- glmer(data=minimal_dataset,
                              curr_formula, family='binomial',
                              control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
    }
    else{ #continuous measures
        filler.model <- lmer(data=minimal_dataset,
                           curr_formula,
                           control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
    }
  
    saveRDS(filler.model, paste0("../data/filler_models/trained_lmers/Filler_", model_name, "_", measure, ".RDS"))

  NULL
}
close(pb)
stopCluster(cl)
```
Fit regression-contingent filler models
gzfw = forward reading time
gzbk = regressive gaze duration
gpbk = regressive go-past time

```{r}
measures <- c('gzfw', 'gzbk', 'gpbk')

num_cores = min(num_multithread_cores, detectCores()-1)
cl <- makeCluster(num_cores)
registerDoSNOW(cl)
all_tasks <- expand.grid(model_name = model_names, measure = measures, stringsAsFactors = FALSE)
total <- nrow(all_tasks)
pb <- txtProgressBar(max = total, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)
result <- 
foreach(i = 1:nrow(all_tasks), .packages=c('dplyr', 'stringr', 'lme4'), .options.snow = opts) %dopar% {
  model_name <- all_tasks$model_name[i]
  full_measure_name <- all_tasks$measure[i]
  measure <- substr(full_measure_name, 1, 2)
  if(model_name == "baseline_surp"){
    curr_formula <- str_replace_all(formula_nosurp, "MEASURE", measure)
    }
  else{
    curr_formula <- str_replace_all(str_replace_all(formula_lm, "MEASURE", measure), "SURPRISAL", model_name)
    }
  vars_in_model <- c(all.vars(as.formula(curr_formula)), 'dataset', 'cond', 'regressed', 'eos_region', 'bos_region')
  if(full_measure_name == 'gzfw'){
    minimal_dataset <- et_with_surprisal %>% select(all_of(vars_in_model)) %>%
                  filter(dataset == "Filler" & !bos_region & !eos_region & !regressed)
  }
  else{
    minimal_dataset <- et_with_surprisal %>% select(all_of(vars_in_model)) %>%
                  filter(dataset == "Filler" & !bos_region & !eos_region & regressed)
  }
  
  filler.model_reg_contingent <- lmer(data=minimal_dataset,
                     curr_formula,
                     control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
  
  saveRDS(filler.model_reg_contingent, paste0("../data/filler_models/trained_lmers/Filler_", model_name, "_", full_measure_name, ".RDS"))
  NULL
  }
close(pb)
stopCluster(cl)
```

Generate predicted reading data from filler models 
```{r}
export_csv <- function(et_data, models, measures, datasets) {
  print('Predicting from filler models:')
  num_cores = min(10, detectCores()-1)
  cl <- makeCluster(num_cores)
  registerDoSNOW(cl)
  all_tasks <- expand.grid(model_name = models, measure = measures, stringsAsFactors = FALSE)
  total <- nrow(all_tasks)
  curr_dataset_count <- 1
  for(curr_dataset in datasets){
    print(paste0('(', curr_dataset_count, '/', length(datasets), ') ', curr_dataset, ' Dataset'))
    pb <- txtProgressBar(max = total, style = 3)
    progress <- function(n) setTxtProgressBar(pb, n)
    opts <- list(progress = progress)
    curr_data <- subset(et_data, dataset == curr_dataset)

    preds_i <- foreach(i = 1:nrow(all_tasks), .packages=c('dplyr', 'stringr', 'lme4'), .combine = cbind, .options.snow = opts) %dopar% {
      model_name <- all_tasks$model_name[i]
      measure <- all_tasks$measure[i]
      filler_model <- readRDS(paste0("../data/filler_models/trained_lmers/Filler_",
                                   model_name, "_",
                                   measure, ".RDS"))
      predicted_i <- predict(filler_model, newdata=curr_data, allow.new.levels = TRUE)
      newcol <- data.frame(predicted_i)
      names(newcol) <- paste0("predicted_", model_name, "_", measure)
      newcol
    }
    close(pb)
    out_fname <- paste0('../data/predicted_reading_data/', curr_dataset, "_predicted.rds")
    print(paste0('Saving to \'', out_fname, '\'...'))
    saveRDS(cbind(curr_data, preds_i), out_fname)
    print('Done!')
    rm(curr_data)
    rm(preds_i)
    curr_dataset_count <- curr_dataset_count + 1
  }
  stopCluster(cl)
}
```

```{r}
model_names <- read.csv('../data/surprisals/all_lm_names.csv')$x
measures <- c('gz', 'gp', 'gzfw', 'gzbk', 'gpbk', 'regout')
datasets <- c("Filler", "ClassicGP", "RelativeClause", 
                     "AttachmentAmbiguity", "Agreement")

export_csv(et_with_surprisal, c(model_names), c(measures), datasets)
```

Get split-half upper-bound delta-log likelihood of filler models:
```{r}
et_with_surprisal <- readRDS("../data/empirical_reading_data/ET_data_processed_w_surprisals.rds")
```


```{r}
set.seed(54321)
total_iterations <- 15
splithalf_dlls <- numeric()
splithalf_num_obs <- numeric()
splithalf_measure_name <- character()
measures <-  c('gz', 'gp', 'regout', 'gzfw', 'gzbk', 'gpbk')
pb = txtProgressBar(min = 0, max = total_iterations*length(measures), initial = 0, style = 3)
stepi <- 0
for(measure in measures){
  reg_contingent <- (measure == 'gzfw' | measure == 'gzbk' | measure == 'gpbk')
  regressed_i <- measure != 'gzfw'
  for(i in 1:total_iterations){
    subject_ids <- unique(et_with_surprisal$subj)
    split_a <- sample(subject_ids, size=round(length(subject_ids)/2), replace=FALSE)
    if(reg_contingent){
      et_with_surprisal_split_a <- subset(et_with_surprisal, (subj %in% split_a) & regressed == regressed_i)
      et_with_surprisal_split_b <- subset(et_with_surprisal, !(subj %in% split_a) & regressed == regressed_i)
    }
    else{
      et_with_surprisal_split_a <- subset(et_with_surprisal, (subj %in% split_a))
      et_with_surprisal_split_b <- subset(et_with_surprisal, !(subj %in% split_a))
    }
    
    
    word_mean_rts_a <- et_with_surprisal_split_a %>% group_by(item, cond, dataset, region) %>%
      summarize(subj_mean_gz = mean(gz, na.rm=T),
                subj_mean_gp = mean(gp, na.rm=T),
                subj_mean_regout = mean(regout, na.rm=T)) %>%
      mutate(subj_mean_gz = (subj_mean_gz-mean(subj_mean_gz))/sd(subj_mean_gz),
             subj_mean_gp = (subj_mean_gp-mean(subj_mean_gp))/sd(subj_mean_gp))
    
    word_mean_rts_b <- et_with_surprisal_split_b %>% group_by(item, cond, dataset, region) %>%
      summarize(subj_mean_gz = mean(gz, na.rm=T),
                subj_mean_gp = mean(gp, na.rm=T),
                subj_mean_regout = mean(regout, na.rm=T)) %>%
      mutate(subj_mean_gz = (subj_mean_gz-mean(subj_mean_gz))/sd(subj_mean_gz),
             subj_mean_gp = (subj_mean_gp-mean(subj_mean_gp))/sd(subj_mean_gp))
    
    et_with_surprisal_split_a <- merge(et_with_surprisal_split_a, word_mean_rts_b, by=c('item', 'cond', 'dataset', 'region'))
    et_with_surprisal_split_b <- merge(et_with_surprisal_split_b, word_mean_rts_a, by=c('item', 'cond', 'dataset', 'region'))
    
    et_with_surprisal_split_a <- et_with_surprisal_split_a %>%
      group_by_at(vars(item, subj, cond)) %>%
          mutate(subj_mean_gz_p1 = lag(subj_mean_gz, order_by = region),
                 subj_mean_gz_p2 = lag(subj_mean_gz_p1, order_by = region),
                 subj_mean_gp_p1 = lag(subj_mean_gp, order_by = region),
                 subj_mean_gp_p2 = lag(subj_mean_gp_p1, order_by = region),
                 subj_mean_regout_p1 = lag(subj_mean_regout, order_by = region),
                 subj_mean_regout_p2 = lag(subj_mean_regout_p1, order_by = region))
    
    et_with_surprisal_split_b <- et_with_surprisal_split_b %>%
      group_by_at(vars(item, subj, cond)) %>%
          mutate(subj_mean_gz_p1 = lag(subj_mean_gz, order_by = region),
                 subj_mean_gz_p2 = lag(subj_mean_gz_p1, order_by = region),
                 subj_mean_gp_p1 = lag(subj_mean_gp, order_by = region),
                 subj_mean_gp_p2 = lag(subj_mean_gp_p1, order_by = region),
                 subj_mean_regout_p1 = lag(subj_mean_regout, order_by = region),
                 subj_mean_regout_p2 = lag(subj_mean_regout_p1, order_by = region))
    
    et_with_surprisal_split_a <- et_with_surprisal_split_a %>% subset(!is.na(logfreq_s) & !is.na(logfreq_p1_s) & !is.na(logfreq_p2_s) & (sent_length != region-1))
    
    et_with_surprisal_split_b <- et_with_surprisal_split_b %>% subset(!is.na(logfreq_s) & !is.na(logfreq_p1_s) & !is.na(logfreq_p2_s) & (sent_length != region-1))
    
    formula_lm <- "MEASURE ~ SURPRISAL + SURPRISAL_p1 + 
      region_s + 
      logfreq_s*length_s + 
      logfreq_p1_s*length_p1_s + 
      (1 + SURPRISAL + SURPRISAL_p1 + logfreq_s + logfreq_p1_s || subj) + 
      (1 | item)"
    
    formula_nosurp <- "MEASURE ~ region_s + 
                     logfreq_s*length_s + 
                     logfreq_p1_s*length_p1_s + 
                     (1 + logfreq_s + logfreq_p1_s || subj) + 
                     (1 | item)"
    
    if(reg_contingent){
      measure_formula <- substr(measure, 1, 2)
    }
    else{
      measure_formula <- measure
    }
    
    model_name <- paste0('subj_mean_', measure_formula)
      
      baseline_formula <- str_replace_all(formula_nosurp, "MEASURE", measure_formula)
      surp_formula <- str_replace_all(str_replace_all(formula_lm, "MEASURE", measure_formula), "SURPRISAL", model_name)
      vars_in_model <- c(all.vars(as.formula(surp_formula)), 'dataset', 'cond', 'eos_region', 'bos_region')
      minimal_dataset_a <- et_with_surprisal_split_a %>% select(all_of(vars_in_model)) %>%
                          filter(dataset == "Filler" & !bos_region & !eos_region)
      minimal_dataset_a <- minimal_dataset_a[complete.cases(minimal_dataset_a),]
      
      minimal_dataset_b <- et_with_surprisal_split_b %>% select(all_of(vars_in_model)) %>%
                          filter(dataset == "Filler" & !bos_region & !eos_region)
      minimal_dataset_b <- minimal_dataset_b[complete.cases(minimal_dataset_b),]
      
      if(measure == 'regout'){
        baseline.model_a <- glmer(data=minimal_dataset_a,
                                    baseline_formula, family='binomial',
                                   control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
        filler.model_a <- glmer(data=minimal_dataset_a,
                              surp_formula, family='binomial',
                              control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
        baseline.model_b <- glmer(data=minimal_dataset_b,
                                    baseline_formula, family='binomial',
                                   control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
        filler.model_b <- glmer(data=minimal_dataset_b,
                              surp_formula, family='binomial',
                              control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
      }
      else{
        baseline.model_a <- lmer(data=minimal_dataset_a,
                               baseline_formula,
                               control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
        
        
        filler.model_a <- lmer(data=minimal_dataset_a,
                               surp_formula,
                               control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
        
        
        baseline.model_b <- lmer(data=minimal_dataset_b,
                               baseline_formula,
                               control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
      
        
        filler.model_b <- lmer(data=minimal_dataset_b,
                               surp_formula,
                               control=lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=2e5)))
      }
      
      #print(summary(filler.model_a))
      #print(summary(filler.model_b))
      dll_a <- logLik(filler.model_a) - logLik(baseline.model_a)
      dll_b <- logLik(filler.model_b) - logLik(baseline.model_b)
      if((nobs(filler.model_b) != nobs(baseline.model_b)) | (nobs(filler.model_a) != nobs(baseline.model_a))){
        print('Error: baseline and surprisal model differ in number of observations')
      }
      writeLines(paste0(measure, ' iteration ', i, ' out of ',  total_iterations, ':', 
                  '\n\tNum. Observations: ', nobs(filler.model_b)+nobs(filler.model_a), 
                  '\n\tTotal DLL: ', dll_a+dll_b,
                  '\n\tPer Obs. DLL: ',(dll_a+dll_b)/((nobs(filler.model_b)+(nobs(filler.model_a))))))
      splithalf_dlls <- c(splithalf_dlls, dll_a+dll_b)
      splithalf_num_obs <- c(splithalf_num_obs, nobs(filler.model_b)+nobs(filler.model_a))
      splithalf_measure_name <- c(splithalf_measure_name, measure)
      stepi <- stepi + 1
      setTxtProgressBar(pb,stepi)
    }
}
close(pb)

upper_bound_dlls <- data.frame(measure=splithalf_measure_name, dll=splithalf_dlls, nobs=splithalf_num_obs)
saveRDS(upper_bound_dlls, '../data/filler_models/upper_bound_dlls.RDS')
```

